---
title: Appendix S1. Detailed instructions for replicating analyses and reproducing
  results.
output:
  html_document:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

***

This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

***

```{r set_options, echo=FALSE, cache=FALSE}
options(width = 100)
knitr::opts_chunk$set(message = FALSE)
```

## Requirements
All analyses require the [R software](https://cran.r-project.org/) (v3.2.3) for data retrieval, data processing, and summarizing model results, and the [JAGS software](http://mcmc-jags.sourceforge.net/) (v4.2.0) for Markov chain Monte Carlo (MCMC) simulation. Please note that some of the R code below may not work with older versions of JAGS due to some changes in the ways that arrays are handled.

We also need a few packages that are not included with the base installation of R, so we begin by installing them (if necessary) and then loading them.

```{r load_pkgs, message=FALSE, warning=FALSE}
if(!require("R2jags")) {
  install.packages("R2jags")
  library("R2jags")
}
if(!require("readr")) {
  install.packages("readr")
  library("readr")
}
if(!require("gsl")) {
  install.packages("gsl")
  library("gsl")
}
```

The last thing we'll need a couple of helper functions.

```{r define_Re2prec}
## better round
Re2prec <- function(x,map="round",prec=1) {
  ## 'map' can be round, floor, or ceiling
  ## 'prec' is nearest value (eg, 0.1 means to nearest tenth); default 1 gives normal behavior
  if(prec<=0) { stop("\"prec\" cannot be less than or equal to 0") }
  do.call(map,list(x/prec))*prec
}

## return spring transition index
get_STI <- function(x, day_max=200) {
	return(min(which(x==min(x[1:day_max]))))
}

## colVars; from Gelman
## returns the column-wise variance of a matrix
colVars <- function(a) {
	n <- dim(a)[[1]]
	c <- dim(a)[[2]]
	mm <- matrix(.colMeans(a, n, c), n, c, byrow = TRUE)
	return(.colMeans(((a - mm) ^ 2), n, c) * n / (n - 1))
}

## waic; from Gelman
## computes WAIC based on pointwise log-like
waic <- function(log_lik) {
  S <- nrow(log_lik)
  n <- ncol(log_lik)
  lpd <- log(colMeans(exp(log_lik)))
  p_waic <- colVars(log_lik)
  elpd_waic <- lpd - p_waic
  waic <- -2*elpd_waic
  loo_weights_raw <- 1/exp(log_lik-max(log_lik))
  loo_weights_normalized <- loo_weights_raw /
    matrix(colMeans(loo_weights_raw),nrow=S,ncol=n,byrow=TRUE)
  loo_weights_regularized <- pmin(loo_weights_normalized, sqrt(S))
  elpd_loo <- log(colMeans(exp(log_lik)*loo_weights_regularized) /
                    colMeans(loo_weights_regularized))
  p_loo <- lpd - elpd_loo
  pointwise <- cbind(waic,lpd,p_waic,elpd_waic,p_loo,elpd_loo)
  total <- colSums(pointwise)
  se <- sqrt(n*colVars(pointwise))
  return(list(waic=total["waic"],
              elpd_waic=total["elpd_waic"],
              p_waic=total["p_waic"],
              elpd_loo=total["elpd_loo"],
              p_loo=total["p_loo"],
              pointwise=pointwise,
              total=total,
              se=se))
}
```

## User inputs
We begin by specifying the names of four necessary data files that contain the following information:
 
 1. observed total number of adult spawners (escapement) by year;
 2. observed age composition of adult spawners by year;
 3. observed total harvest by year;
 4. hatchery releases by year.

Let's also define the following parameters, which will be referenced throughout the analysis.

 * `n_yrs`: number of calendar years of data
 * `A`: number of age classes 
 * `M`: number of covariates

```{r get_user_inputs}
## 1. file with escapement data
## [n_yrs x 2] matrix of obs counts; 1st col is calendar yr
fn_esc <- "skagit_sthd_esc.csv"

## 2. file with age comp data
## [n_yrs x (1+A)]; 1st col is calendar yr
fn_age <- "skagit_sthd_age.csv"
## min & max ages
age_min <- 3
age_max <- 8
## years, if any, of age-comp to skip; see below
age_skip <- 2

## 3. file with harvest data
## [n_yrs x 2] matrix of obs catch; 1st col is calendar yr
fn_harv <- "skagit_sthd_catch.csv"

## 4. file with hatchery release data
## [n_yrs x (1+MM)]; 1st col is calendar yr
fn_hrel <- "skagit_sthd_hrel.csv"

## time lags (years) for covariates
flow_lag <- 1
marine_lag <- 2
hrel_lag <- 2

## number of years of forecasts
n_fore <- 1

## file where to save JAGS model
fn_jags <- "skagit_sthd_jags.txt"

## upper threshold for Gelman & Rubin's (1992) potential scale reduction factor (Rhat).
Rhat_thresh <- 1.1

## URL for example data files
## set to NULL if using a local folder/directory
ex_url <- "https://raw.githubusercontent.com/mdscheuerell/skagit_sthd/master/data/"
```

## Loading the fish data
Here we load in the first three data files and do some simple calculations and manipulations. First the spawner data:

```{r get_escapement_data}
## escapement
dat_esc <- read_csv(paste0(ex_url,fn_esc))
## years of data
dat_yrs <- dat_esc$year
## number of years of data
n_yrs <- length(dat_yrs)
## get first & last years
yr_frst <- min(dat_yrs)
yr_last <- max(dat_yrs)
## log of escapement
ln_dat_esc <- c(log(dat_esc$escapement),rep(NA,n_fore))
```

Next the age composition data:

```{r get_age_data}
## age comp data
dat_age <- read_csv(paste0(ex_url,fn_age))
## drop year col & first age_min+age_skip rows
dat_age <- dat_age[-(1:(age_min+age_skip)),-1]
## num of age classes
A <- age_max-age_min+1
## add row(s) of NA's for forecast years
dat_age <- rbind(dat_age,matrix(0,n_fore,A,dimnames=list(n_yrs+seq(n_fore),colnames(dat_age))))
## total num of age obs by cal yr
dat_age[,"sum"] <- apply(dat_age,1,sum)
## row indices for any years with no obs age comp
idx_NA_yrs <- which(dat_age$sum<A,TRUE)
## replace 0's in yrs w/o any obs with NA's
dat_age[idx_NA_yrs,(1:A)] <- NA
## change total in yrs w/o any obs from 0 to A to help dmulti()
dat_age[idx_NA_yrs,"sum"] <- A
## convert class
dat_age <- as.matrix(dat_age)
```

And then the harvest data:

```{r get_harvest}
## harvest
dat_harv <- read_csv(paste0(ex_url,fn_harv))
## drop year col & first age_max rows
dat_harv <- c(dat_harv$catch,rep(0,n_fore))
```

## Loading the covariates

Here we use a dummy covariate, which is simply a vector of all 0's.

```{r combine_covars}
## transform the covariates to z-scores
mod_cvrs <- rep(0, n_yrs-age_min+n_fore) 
## total number of covariates
n_cov <- 1 
```

## Specifying model in JAGS

Now we can specify the model in JAGS. Note that the code below is not written to be universally generic with respect to the number of covariates, but rather to emphasize how to incorporate the three we used in this specific application. The important thing is the number of covariate parameters in the `PRIORS` and `LIKELIHOOD` sections (i.e., there must be a unique `c_Name` parameter for each of the _MM_ covariates).

```{r SSSR_in_JAGS}
cat("

model {
	
	##--------
	## PRIORS
	##--------
	## alpha = exp(a) = intrinsic productivity
	alpha ~ dunif(0.1,10);
	mu_BH_a <- log(alpha);
	E_BH_a <- mu_BH_a + sigma_r/(2 - 2*phi^2);

	## strength of dens depend
	beta ~ dunif(0,0.01);

	## AR(1) coef for proc errors
	phi ~ dunif(-0.999,0.999);
	
	## process variance for recruits model
	sd_r ~ dunif(0.001,20);
	tau_r <- pow(sd_r,-2);
	sigma_r <- pow(sd_r,2);
	
	## innovation in first year
	innov_1 ~ dnorm(0,tau_r*(1-phi*phi));
	
	## obs variance for spawners
	sd_s ~ dunif(0.001,20);
	tau_s <- pow(sd_s,-2);
	sigma_s <- pow(sd_s,2);
	
	## unprojectable early recruits;
	## hyper mean across all popns
	Rec_mu ~ dnorm(0,0.001);
	## hyper SD across all popns
	Rec_sig ~ dunif(0,100);
	## precision across all popns
	Rec_tau <- pow(Rec_sig,-2);
	## multipliers for unobservable total runs
	ttl_run_mu ~ dunif(1,5);
	ttl_run_tau ~ dunif(1,20);
	
	## maturity schedule
	## unif vec for Dirch prior
	for(i in 1:A) { theta[i] <- 1 }
	## hyper-mean for maturity
	pi_eta ~ ddirch(theta);
	## hyper-prec for maturity
	pi_tau ~ dunif(0.001,1e3);
	for(t in 1:(n_yrs-age_min+n_fore)) { pi_vec[t,1:A] ~ ddirch(pi_eta*pi_tau) }
	
	##------------
	## LIKELIHOOD
	##------------
	## 1st brood yr requires different innovation
	## predicted recruits in BY t
  ln_BH_a[1] <- mu_BH_a; 
	E_ln_Rec[1] <- ln_BH_a[1] + ln_Sp[1] - log(1 + beta*Sp[1]) + phi*innov_1;
	tot_ln_Rec[1] ~ dnorm(E_ln_Rec[1],tau_r);
	res_ln_Rec[1] <- tot_ln_Rec[1] - E_ln_Rec[1];
	## median of total recruits
	tot_Rec[1] <- exp(tot_ln_Rec[1]);

	## R/S
	ln_RS[1] <- tot_ln_Rec[1] - ln_Sp[1];
		
	## brood-yr recruits by age
	for(a in 1:A) {
		Rec[1,a] <- max(1,tot_Rec[1] * pi_vec[1,a]);
	}
	
	## brood years 2:(n_yrs-age_min)
	for(t in 2:(n_yrs-age_min+n_fore)) {
		## predicted recruits in BY t
    ln_BH_a[t] <- mu_BH_a; 
	  E_ln_Rec[t] <- ln_BH_a[t] + ln_Sp[t] - log(1 + beta*Sp[t]) + phi*res_ln_Rec[t-1];
		tot_ln_Rec[t] ~ dnorm(E_ln_Rec[t],tau_r);
		res_ln_Rec[t] <- tot_ln_Rec[t] - E_ln_Rec[t];
		## median of total recruits
		tot_Rec[t] <- exp(tot_ln_Rec[t]);
		## R/S
		ln_RS[t] <- tot_ln_Rec[t] - ln_Sp[t];
		## brood-yr recruits by age
		for(a in 1:A) {
			Rec[t,a] <- max(1,tot_Rec[t] * pi_vec[t,a]);
		}
	} ## end t loop over year

	## get total cal yr returns for first age_min yrs
	for(i in 1:(age_min+age_skip)) {
		ln_tot_Run[i] ~ dnorm(ttl_run_mu*Rec_mu,Rec_tau/ttl_run_tau);
		tot_Run[i] <- exp(ln_tot_Run[i]);
	}

	## get predicted calendar year returns by age
	## matrix Run has dim [(n_yrs-age_min) x A]
	## step 1: incomplete early broods
	## first cal yr of this grp is first brood yr + age_min + age_skip
	for(i in 1:(age_max-age_min-age_skip)) {
		## projected recruits
		for(a in 1:(i+age_skip)) {
			Run[i,a] <- Rec[(age_skip+i)-a+1,a];
		}
		## imputed recruits
		for(a in (i+1+age_skip):A) {
			lnRec[i,a] ~ dnorm(Rec_mu,Rec_tau);
			Run[i,a] <- exp(lnRec[i,a]);
		}
		## total run size
		tot_Run[i+age_min+age_skip] <- sum(Run[i,1:A]);
		# predicted age-prop vec for multinom
		for(a in 1:A) {
			age_v[i,a] <- Run[i,a] / tot_Run[i+age_min];
		}
		## multinomial for age comp
		dat_age[i,1:A] ~ dmulti(age_v[i,1:A],dat_age[i,A+1]);
		lp_age[i] <- logdensity.multi(dat_age[i,1:A],age_v[i,1:A],dat_age[i,A+1]);
	}
	
	## step 2: info from complete broods
	## first cal yr of this grp is first brood yr + age_max
	for(i in (A-age_skip):(n_yrs-age_min-age_skip+n_fore)) {
		for(a in 1:A) {
			Run[i,a] <- Rec[(age_skip+i)-a+1,a];
		}
		## total run size
		tot_Run[i+age_min+age_skip] <- sum(Run[i,1:A]);
		## predicted age-prop vec for multinom
		for(a in 1:A) {
			age_v[i,a] <- Run[i,a] / tot_Run[i+age_min];
		}
		## multinomial for age comp
		dat_age[i,1:A] ~ dmulti(age_v[i,1:A],dat_age[i,A+1]);
		lp_age[i] <- logdensity.multi(dat_age[i,1:A],age_v[i,1:A],dat_age[i,A+1]);
	}
		
	## get predicted calendar year spawners
	## first cal yr is first brood yr
	for(t in 1:(n_yrs+n_fore)) {
		## obs model for spawners
		Sp[t] <- max(1,tot_Run[t] - dat_harv[t]);
		ln_Sp[t] <- log(Sp[t]);
		ln_dat_esc[t] ~ dnorm(ln_Sp[t], tau_s);
		lp_esc[t] <- logdensity.norm(ln_dat_esc[t],ln_Sp[t], tau_s);
	}
			
} ## end model description

", file=fn_jags)
```

***
## Fitting the model

The last thing we need to do before fitting the model in JAGS is to specify:

1. the data and indices that go into the model;
2. the model parameters and states that we want JAGS to return;
3. the MCMC control parameters.

Please note that the following code takes ~20 min to run on a quad-core machine with 3.5 GHz Intel processors.

```{r start_timer, include=FALSE}
## start timer
timer_start <- proc.time() 
```

```{r JAGS_IO, message=FALSE, warning=FALSE, cache=TRUE}
## data to pass to JAGS
dat_jags <- c("dat_age","ln_dat_esc","dat_harv","mod_cvrs","n_cov",
              "n_yrs","A","age_min","age_max","age_skip","n_fore") 

## 2. model params/states for JAGS to return
#par_jags <- c("alpha","E_BH_a","mu_BH_a","beta","ln_BH_a",
#              "lp_age","lp_esc","phi",
#              "Sp","Rec","tot_ln_Rec","ln_RS","pi_vec",
#              "sigma_r","sigma_s","res_ln_Rec")
par_jags <- c("lp_age","lp_esc")

## 3. MCMC control params
## MCMC parameters
mcmc_chains <- 4
mcmc_length <- 5e4
mcmc_burn <- 2.5e4
mcmc_thin <- 50
## total number of MCMC samples
mcmc_samp <- (mcmc_length-mcmc_burn)*mcmc_chains/mcmc_thin

## function to create JAGS inits
init_vals <- function() {
	list(alpha=2,
	     beta=1/exp(mean(ln_dat_esc, na.rm=TRUE)),
	     pi_tau=1, pi_eta=rep(1,A),
	     pi_vec=matrix(c(0.01,0.3,0.48,0.15,0.05,0.01),n_yrs-age_min+n_fore,A,byrow=TRUE),
	     Rec_mu=log(1000),
	     Rec_sig=0.1,
	     tot_ln_Rec=rep(log(1000),n_yrs-age_min+n_fore),
	     innov_1=0,
	     phi=0.5)
	}

## list of model info for JAGS
mod_jags <- list(data=dat_jags,
                 inits=init_vals,
                 parameters.to.save=par_jags,
                 model.file=fn_jags,
                 n.chains=as.integer(mcmc_chains),
                 n.iter=as.integer(mcmc_length),
                 n.burnin=as.integer(mcmc_burn),
                 n.thin=as.integer(mcmc_thin),
                 DIC=TRUE)

## fit the model in JAGS & store results
rk_fits <- do.call(jags.parallel, mod_jags)
```

```{r stop_timer, include=FALSE}
## stop timer
run_time_in_min <- round(((proc.time()-timer_start)/60)["elapsed"], 1)
cat(run_time_in_min, file="run_time_in_min.txt")
```

## Finding the best model

Let's examine Watanabe's Akaike Information Criterion (WAIC) for each of the models to see which of the covariate scenarios seems to be best supported by the available data.

```{r get_WAIC}
## extract log densities from JAGS objects
ldens <- cbind(rk_fits$BUGSoutput$sims.list$lp_age,
               rk_fits$BUGSoutput$sims.list$lp_esc)
(WAIC <- waic(ldens)$waic)
(LOO <- -2*waic(ldens)$elpd_loo)
```

