---
title: Appendix S3. Steps to recreate figures from main text.
subtitle: Supporting information for Scheuerell et al.
output:
  pdf_document:
    highlight: haddock
    toc: yes
    number_sections: true
    toc_depth: '3'
fontsize: 11pt
geometry: margin=1in
---

```{r set_options, echo = FALSE, message = FALSE}
options(width = 100)
knitr::opts_chunk$set(message = FALSE)
```

\vspace{0.2in}

This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

# Background

This appendix shows how to recreate the figures in the main text based on the results from the best of the fitted models.

All analyses require the [R software](https://cran.r-project.org/) (v3.4.3 or later), as well as a few packages that are not included with the base installation of R.

```{r load_pkgs, message = FALSE, warning = FALSE}
if(!require("readr")) {
  install.packages("readr")
  library("readr")
}
if(!require("coda")) {
  install.packages("coda")
  library("coda")
}
if(!require("here")) {
  install.packages("here")
  library("here")
}
## set directory locations
datadir <- here("data")
analdir <- here("analysis")
savedir <- here("analysis/cache")
```

We also need a couple of helper functions.

```{r define_funcs}
## better round
Re2prec <- function(x, map = "round", prec = 1) {
  ## 'map' can be "round", "floor", or "ceiling"
  ## 'prec' is nearest value (eg, 0.1 means to nearest tenth; 1 gives normal behavior)
  if(prec<=0) { stop("\"prec\" cannot be less than or equal to 0") }
  do.call(map,list(x/prec))*prec
}
```

```{r get_user_inputs, echo=FALSE}
## first & last years of fish data
yr_frst <- 1978
yr_last <- 2018
## years of data
dat_yrs <- seq(yr_frst,yr_last)
## number of years of data
n_yrs <- length(dat_yrs)

## min & max adult age classes
age_min <- 3
age_max <- 8
## num of age classes
A <- age_max - age_min + 1

## number of years for run forecasts
n_fore <- 0

## Posterior coverage interval
CI_vec <- c(0.025,0.5,0.975)
```

# Load the information

Here we load in the model fits, covariates, and harvest data.

```{r load_mod_fits}
best_fit <- readRDS(file.path(savedir, "fit_bh_cov.rds"))
```

```{r load_covariates}
## covariate(s)
dat_cvrs <- read_csv(file.path(datadir, "skagit_sthd_covars.csv"))
## total number of covariates
n_cov <- dim(dat_cvrs)[2] - 1
```

```{r get_escapement_data}
## escapement
dat_esc <- read_csv(file.path(datadir, "skagit_sthd_esc.csv"))
## log of escapement
ln_dat_esc <- c(log(dat_esc$escapement), rep(NA, n_fore))
```

```{r get_harvest}
## harvest
dat_harv <- read_csv(file.path(datadir, "skagit_sthd_catch.csv"))
## drop year col & first age_max rows
dat_harv <- c(dat_harv$catch, rep(0, n_fore))
```

# Main results

Here is a table of summary statistics for some of the model parameters.

```{r tbl_summary_stats}
## params of interest
par_conv <- c("alpha", "beta", paste0("gamma[",seq(n_cov),"]"),
              "sigma_r", "sigma_s",
              "pi_tau", paste0("pi_eta[",seq(A-1),"]"))
## summary table
tbl_smry <- summary(best_fit)$quantiles[par_conv, c("2.5%","50%","97.5%")]
round(tbl_smry, 2)
```

Now we need to convert the `mcmc.list` output into a more user-friendly form for plotting, etc.

```{r}
## results
mod_res <- do.call("rbind", best_fit)
```

## Total population size

Here is our estimate of the total run size (i.e., catch + escapement) over time, which includes a forecast for `r yr_last+n_fore`. The black points are the data, the blue line is the median posterior estimate, and the shaded region is the 95% credible interval. Note that the y-axis is on a log scale.

```{r plot_run_size, fig.width = 6, fig.height = 4.5, fig.pos = "placeHere"}
clr <- rgb(0, 0, 255, alpha = 50, maxColorValue = 255)
## estimated spawner data for plotting
p_dat <- mod_res[,grep("Sp", colnames(mod_res))]
p_dat <- apply(p_dat, 2, quantile, CI_vec)
p_dat <- p_dat + matrix(dat_harv, length(CI_vec), n_yrs+n_fore, byrow = TRUE)
## time seq
t_idx_f <- seq(yr_frst, length.out = n_yrs+n_fore)
## plot
yp_min <- min(p_dat)
yp_max <- max(p_dat)
par(mai = c(0.8,0.8,0.1,0.1), omi = c(0.5,0.2,0.1,0.2))
plot(t_idx_f, p_dat[3,], ylim = c(yp_min,yp_max), type = "n",
     log = "y", xaxt = "n", yaxt = "n", bty = "L",
     xlab = "Year", ylab = "Catch + escapement", main = "", cex.lab = 1.2)
polygon(c(t_idx_f, rev(t_idx_f)), c(p_dat[3,], rev(p_dat[1,])),
        col = clr, border = NA)
lines(t_idx_f, p_dat[2,], col = "blue3", lwd = 2)
points(t_idx_f, exp(ln_dat_esc)+dat_harv, pch = 16, cex = 1)
axis(1, at = seq(1980, 2015, 5))
axis(2, at = c(4000, 8000, 16000))
```

## Spawner-recruit relationship

Here is the relationship between spawner and subsequent recruits (a), assuming mean values for all covariates. Gray lines show the median relationship for each of the `r n_yrs` years based on $a_t$. Note that for plotting purposes only in (b) and (c), the density in the largest bin for each parameter contains counts for all values greater or equal to that. Vertical arrows under the x-axes in (b) and (c) indicate the 2.5^th^, 50^th^, and 97.5^th^ percentiles.

```{r plot_S_R, fig.width = 8, fig.height = 5, fig.pos = "placeHere"}
layout(matrix(c(1,1,2,3),2,2),c(3,2),c(1,1))
xoffSet <- 0.05
yoffSet <- 0.03

## colors for plotting
clr <- rgb(100, 0, 200,
           alpha = seq(200, 100,
                       length.out = age_max-age_min+n_fore),
           maxColorValue = 255)

## posterior of spawners
s_dat <- mod_res[,grep("Sp", colnames(mod_res))]
s_dat <- apply(s_dat, 2, quantile, CI_vec)
s_dat <- s_dat[, 1:(n_yrs-age_min+n_fore)]

## posterior of recruits
r_dat <- mod_res[, grep("tot_ln_Rec", colnames(mod_res))]
r_dat <- exp(apply(r_dat, 2, quantile, CI_vec))

## median values for a & b
aa <- apply(mod_res[, grep("ln_BH_a", colnames(mod_res))], 2, median)
bb <- median(mod_res[, "beta"])

## empty plot space for spawner-recruit relationships
dd <- 3000
yM <- Re2prec(max(r_dat), "ceiling", dd)
xM <- Re2prec(max(s_dat), "ceiling", dd)
par(mai = c(0.8,0.8,0.1,0.1), omi = c(0,0,0,0))
plot(s_dat[2,], r_dat[2,], xlim = c(0,xM), ylim = c(0,yM), type = "n",
     xaxs = "i", yaxs = "i", cex.lab = 1.2,
     xlab = expression(Spawners~(10^3)),
     ylab = expression(Recruits~(10^3)),
     xaxt = "n", yaxt = "n", bty="L")
axis(1, at = seq(0,xM,dd*2), labels = seq(0,xM,dd*2)/1000)
axis(2, at = seq(0,yM,dd*2), labels = seq(0,yM,dd*2)/1000, las=1)
for(i in 1:length(aa)) {
  lines(exp(aa[i]) * seq(0,xM) / (1 + bb * seq(0,xM)),
        col = "darkgray")
  }
abline(a = 0,b = 1,lty = "dashed")

## add S-R estimates and medians
nCB <- n_yrs-age_max
## years with complete returns
points(s_dat[2, 1:nCB], r_dat[2, 1:nCB],
       xlim = c(0,xM), ylim = c(0,yM),
       pch = 16, col = "blue3")
segments(s_dat[2, 1:nCB], r_dat[1, 1:nCB],
         s_dat[2, 1:nCB], r_dat[3, 1:nCB],
         col = "blue3")
segments(s_dat[1, 1:nCB], r_dat[2, 1:nCB],
         s_dat[3, 1:nCB], r_dat[2, 1:nCB],
         col = "blue3")
nTB <- dim(s_dat)[2]
## years with incomplete returns
segments(s_dat[2, (nCB+1):nTB], r_dat[1, (nCB+1):nTB],
         s_dat[2, (nCB+1):nTB], r_dat[3, (nCB+1):nTB],
         col = clr)
segments(s_dat[1, (nCB+1):nTB], r_dat[2, (nCB+1):nTB],
         s_dat[3, (nCB+1):nTB], r_dat[2, (nCB+1):nTB],
         col = clr)
points(s_dat[2, (nCB+1):nTB],r_dat[2, (nCB+1):nTB],
       xlim = c(0,xM), ylim = c(0,yM),
       pch = 16, col = clr)
text(x = par()$usr[1] + diff(par()$usr[1:2]) * xoffSet,
     y = par()$usr[4] - diff(par()$usr[3:4]) * yoffSet,
     "(a)")

## posterior for alpha
clr <- rgb(0, 0, 255, alpha = 50, maxColorValue = 255)
a_thresh <- 59
par(mai = c(0.8,0.4,0.3,0.1))
## B-H alpha
R_alpha_est <- mod_res[, "alpha"]
alphaCI <- quantile(R_alpha_est, CI_vec)
R_alpha_est[R_alpha_est > a_thresh] <- a_thresh
hist(R_alpha_est, freq = FALSE, breaks = seq(0, a_thresh+1, 2),
     col = clr, border = "blue3",
     xlab = "", ylab = "", main = "", cex.lab = 1.2, yaxt = "n")
aHt <- (par()$usr[4]-par()$usr[3])/12
arrows(alphaCI, par()$usr[3], alphaCI,par()$usr[3]-aHt,
       code = 1, length = 0.05, xpd = NA, col = "blue3", lwd = 1.5)
mtext(expression(Instrinsic~productivity~(alpha)), 1, line = 3, cex = 1)
text(x = par()$usr[1],
     y = par()$usr[4] * 1.05,
     "(b)", xpd=NA)

## posterior for K
par(mai = c(0.8,0.4,0.3,0.1))
aa <- mod_res[, "alpha"]
bb <- mod_res[, "beta"]
## K in 1000s
R_b_est <- (aa-1) / bb / 1000
R_b_est <- R_b_est[R_b_est > 0]
R_b_CI <- quantile(R_b_est, CI_vec)
## pile into last ban for plotting
R_b_est[R_b_est > 13] <- 13
brks <- seq(Re2prec(min(R_b_est), "floor"),
            Re2prec(max(R_b_est), "ceiling"),
            length.out = length(seq(0, a_thresh, 2)))
hist(R_b_est, freq = FALSE, breaks = brks, col = clr, border = "blue3",
     xlab = "", xaxt = "n", yaxt = "n",
     main = "", ylab = "", cex.lab = 1.2)
axis(1, at = seq(Re2prec(min(R_b_est), "floor"),
                 Re2prec(max(R_b_est), "ceiling"),
                 2))
aHt <- (par()$usr[4] - par()$usr[3]) / 12
arrows(R_b_CI, par()$usr[3], R_b_CI,par()$usr[3]-aHt,
       code = 1, length = 0.05, xpd = NA, col = "blue3", lwd = 1.5)
mtext(expression(paste("Carrying capacity (",italic(K),", ",10^3,")")),
      side = 1, line = 3, cex = 1)
text(x = par()$usr[1], 
     y = par()$usr[4] * 1.05,
     "(c)", xpd=NA)
```


## Covariate effects

Here are time series plots of the covariates (a-c) and histograms of their effects on productivity (d-f).

```{r plot_cov_effects, fig.width=8, fig.height=10, fig.pos="placeHere", warnings=FALSE, messages=FALSE}
clr <- rgb(0, 0, 255, alpha = 50, maxColorValue = 255)
xoffSet <- 0.04
yoffSet <- 0.03
c_est <- mod_res[,grep("gamma", colnames(mod_res))]

par(mfrow=c(n_cov,2), mai=c(0.4,0.2,0.1,0.1), omi=c(0.2,0.5,0,0))
ylN <- floor(min(c_est)*10)/10
ylM <- ceiling(max(c_est)*10)/10
brks <- seq(ylN,ylM,length.out=diff(c(ylN,ylM))*40+1)
cov_names <- c(expression(paste("Max flow (",m^3," ",s^{-1},")")),
               expression(paste("Min flow (",m^3," ",s^{-1},")")),
               "NPGO",
               expression(paste("H releases (",10^3,")")))
t_idx <- seq(yr_frst,length.out=n_yrs-age_min+n_fore)
dat_cvrs <- as.matrix(dat_cvrs[seq(length(t_idx)),])

for(i in 1:n_cov) {
  if(i==4) {
    dat_cvrs[,i+1] <- dat_cvrs[,i+1]/1000
  }
  ## plot covar ts
  plot(dat_cvrs[,"year"], dat_cvrs[,i+1],
       pch=16, col="blue3", type="o",
       xlab="", ylab="", main="", bty = "L",
       cex.axis=1.2)
  text(x = par()$usr[1] + diff(par()$usr[1:2]) * xoffSet,
       y = par()$usr[4] - diff(par()$usr[3:4]) * yoffSet,
       paste0("(",letters[i],")"),
       cex=1.2)
  mtext(side=2, cov_names[i], line=3, cex=1.2)
  if(i==n_cov) { mtext(side=1,"Brood year", line=3) }
  ## plot covar effect
  hist(c_est[,i],
       freq=FALSE,breaks=brks,col=clr,border="blue3",
       xlab="", yaxt="n", main="", ylab="", cex.axis=1.2)
  c_CI <- quantile(c_est[,i],CI_vec)
  aHt <- (par()$usr[4]-par()$usr[3])/20
  arrows(c_CI,par()$usr[3]-0.005,c_CI,par()$usr[3]-aHt,
         code=1,length=0.05,xpd=NA,col="blue3",lwd=1.5)
  abline(v=0, lty="dashed")
  text(x = par()$usr[1] + diff(par()$usr[1:2]) * xoffSet,
       y = par()$usr[4] - diff(par()$usr[3:4]) * yoffSet,
       paste0("(",letters[i+n_cov],")"),
       cex=1.2)
  if(i==n_cov) { mtext(side=1,"Effect size", line=3) }
}
```


## Process errors

Here is the time series of the residuals from the process model. They represent the population's productivity after accounting for the effects of density dependence and environmental covariates.

```{r plot_proc_err, fig.width = 6, fig.height = 4, fig.pos = "placeHere"}
## time sequence
t_idx_a <- seq(yr_frst, length.out = n_yrs-age_min+n_fore)
## plot data
p_dat <- mod_res[,grep("res_ln_Rec", colnames(mod_res))]
p_dat <- apply(p_dat, 2, quantile, CI_vec)
yp_min <- min(p_dat)
yp_max <- max(p_dat)
## plot
par(mai = c(0.8,0.8,0.1,0.1), omi = c(0,0.2,0.1,0.2))
plot(t_idx_a,p_dat[3,], ylim = c(yp_min,yp_max), type = "n",  bty = "L",
     xlab = "Brood year", ylab = "Process error", main = "", cex.lab = 1.2)
abline(h = 0, lty = "dashed")
polygon(c(t_idx_a,rev(t_idx_a)),c(p_dat[3,],rev(p_dat[1,])),
        col = clr, border = NA)
lines(t_idx_a, p_dat[2,], col = "blue3", lwd = 2)
```
