---
  title: Appendix S2. Model fitting and evaluation
output:
  html_document:
  fig_caption: yes
theme: cerulean
toc: yes
toc_depth: 3
toc_float: yes
number_sections: true
pdf_document:
  toc: yes
toc_depth: '3'
---
  
  ***
  
  This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

***
  
  ```{r set_options, echo = FALSE, message = FALSE}
options(width = 100)
knitr::opts_chunk$set(message = FALSE)
set.seed(123)
if(file.exists("cnt_time.txt")) {
  file.remove("cnt_time.txt")
}
## setup parallel processing
options(mc.cores = parallel::detectCores())
```

# Requirements

All analyses require the [R software](https://cran.r-project.org/) (v3.4.3) for data retrieval, data processing, and summarizing model results, and the [Stan software](http://mc-stan.org/) (v2.17) for Hamiltonian Monte Carlo (HMC) simulation.

We also need a few packages that are not included with the base installation of R, so we begin by installing them (if necessary) and then loading them.

```{r load_pkgs, message = FALSE, warning = FALSE}
if(!require("here")) {
  install.packages("here")
  library("here")
}
if(!require("readr")) {
  install.packages("readr")
  library("readr")
}
# (install if necessary &) load pkgs
if(!require("coda")) {
  install.packages("coda")
  library("coda")
}
if(!require("gtools")) {
  install.packages("gtools")
  library("gtools")
}
if(!require("rstan")) {
  install.packages("rstan")
  library("rstan")
}
if(!require("shinystan")) {
  install.packages("shinystan")
  library("shinystan")
}
if(!require("loo")) {
  install.packages("loo")
  library("loo")
}
if(!require("ggplot2")) {
  install.packages("ggplot2")
  library("ggplot2")
}
## set directory locations
datadir <- here("data")
standir <- here("stan")
analdir <- here("analysis")
## write Stan files to disk
rstan_options(auto_write = TRUE)
```

We also need a couple of helper functions.

```{r define_funcs}
## better round
Re2prec <- function(x, map = "round", prec = 1) {
  ## 'map' can be "round", "floor", or "ceiling"
  ## 'prec' is nearest value (eg, 0.1 means to nearest tenth; 1 gives normal behavior)
  if(prec<=0) { stop("\"prec\" cannot be less than or equal to 0") }
  do.call(map,list(x/prec))*prec
}

## convert Stan object to coda
stan2coda <- function(fit) 
{
  mcmc.list(lapply(1:ncol(fit), function(x) mcmc(as.array(fit)[,x,])))
}

## count number of divergent transitions
count_divergences <- function(fit) { 
  sampler_params <- get_sampler_params(fit, inc_warmup=FALSE) 
  sum(sapply(sampler_params, function(x) c(x[,'n_divergent__']))[,1]) 
} 

## histogram of sampler tree depth
hist_treedepth <- function(fit, plot.hist=TRUE) { 
  sampler_params <- get_sampler_params(fit, inc_warmup=FALSE) 
  hist(sapply(sampler_params,function(x) c(x[,'treedepth__']))[,1],breaks=0:20,main="",xlab="Treedepth",plot=plot.hist) 
  abline(v=10, col=2, lty=1) 
} 

## better unlist
unlist2 <- function(x,rr=TRUE,nn=TRUE) {
  if(length(x)==1) {x[[1]]} else {unlist(x,recursive=rr,use.names=nn)}
}
```

# User inputs

We begin by supplying values for the following parameters, which we need for model fitting and evaluation.

```{r get_user_inputs}
## first & last years of fish data
yr_frst <- 1978
yr_last <- 2017

## min & max adult age classes
age_min <- 3
age_max <- 8
## years (if any) of age-comp to skip; see below
age_skip <- 0

## number of years for run forecasts
n_fore <- 0

## upper threshold for Gelman & Rubin's potential scale reduction factor (Rhat).
Rhat_thresh <- 1.1
```

Next we specify the names of three necessary data files containing the following information:
  
  1. observed total number of adult spawners (escapement) by year;
2. observed age composition of adult spawners by year;
3. observed total harvest by year;

```{r get_filenames}
## 1. file with escapement data
## [n_yrs x 2] matrix of obs counts; 1st col is calendar yr
fn_esc <- "skagit_sthd_esc.csv"

## 2. file with age comp data
## [n_yrs x (1+AA)]; 1st col is calendar yr
fn_age <- "skagit_sthd_age.csv"

## 3. file with harvest data
## [n_yrs x 2] matrix of obs catch; 1st col is calendar yr
fn_harv <- "skagit_sthd_catch.csv"
```

# Loading the fish data
Here we load in the first three data files and do some simple calculations and manipulations. First the spawner data:
  
  ```{r get_escapement_data}
## escapement
dat_esc <- read_csv(file.path(datadir, fn_esc))
## years of data
dat_yrs <- dat_esc$year
## number of years of data
n_yrs <- length(dat_yrs)
## drop year col
dat_esc <- dat_esc$escapement
## log of escapement
# ln_dat_esc <- c(log(dat_esc$escapement),rep(NA,n_fore))
## years with missing data
yr_NA_esc <- is.na(dat_esc)
```

Next the age composition data:
  
  ```{r get_age_data}
## age comp data
dat_age <- read_csv(file.path(datadir, fn_age))
## drop year col & first (age_min + age_skip) rows
dat_age <- dat_age[-(1:(age_min+age_skip)),-1]
## num of age classes
AA <- age_max - age_min + 1
## add row(s) of NA's for forecast years
if(n_fore > 0) {
  dat_age <- rbind(dat_age,
                   matrix(0, n_fore, AA,
                          dimnames = list(n_yrs+seq(n_fore),colnames(dat_age))))
}
## total num of age obs by cal yr
dat_age[,"sum"] <- apply(dat_age, 1, sum)
## row indices for any years with no obs age comp
idx_NA_yrs <- which(dat_age$sum == 0, TRUE)
## replace 0's in yrs w/o any obs with NA's
# dat_age[idx_NA_yrs,(1:AA)] <- NA
## change total in yrs w/o any obs from 0 to AA to help dmulti()
dat_age[idx_NA_yrs,"sum"] <- AA
## convert class
dat_age <- as.matrix(dat_age[,1:AA])
```

And then the harvest data:
  
  ```{r get_harvest}
## harvest
dat_harv <- read_csv(file.path(datadir, fn_harv))
## drop year col & first age_max rows
dat_harv <- c(dat_harv$catch, rep(0,n_fore))
```

# Loading the covariates

Our analysis investigates the effects of 4 covariates on the population's instrinic growth rate:

1. Maximum river discharge in winter;
2. Minimum river discharge in summer;
3. North Pacific Gyre Oscillation;
4. Releases of hatchery-born juveniles.

All of the covariates are contained in the file `/data/skagit_sthd_covars.csv`. We will load and then standardize them to have zero-mean and unit-variance.

```{r get_covariates}
## covariate(s)
dat_cvrs <- read_csv(file.path(datadir, "skagit_sthd_covars.csv"))
## drop year col
dat_cvrs <- dat_cvrs[,-1] 
## transform the covariates to z-scores
scl_cvrs <- as.matrix(scale(dat_cvrs)) 
## total number of covariates
n_cov <- dim(scl_cvrs)[2] 
```

# Specifying models in Stan

Now we can specify the various models in Stan. We fit a total of 4 different models, which we outline below, based on the 2 different process models with and without and covariates.

## Ricker with AR(1)

```{r stan_RK_AR}
cat("
functions {
// spawner-recruit functions
real ricker(real log_a, real b, real S) {
real R;
R = S * exp(log_a - S * b);
return(R);
}
}

data {
// number of years of spawner observations
int<lower=1> n_yrs;                       
// number of adult age classes
int<lower=1> AA;                       
// minimum adult age
int<lower=1> age_min;          
// maximum adult age
int<lower=age_min> age_max;
// early years to skip
int<lower=0> age_skip;
// number of years to forecast
int<lower=0> n_fore;          
// number of missing recruits to impute
int<lower=1> n_rec_missing;
// number of years with missing spawner abundance obs
int<lower=0,upper=n_yrs> n_S_NA;          
// rows with missing spawner abundance obs
int<lower=1,upper=n_yrs> which_S_NA[max(n_S_NA,1)]; 
// observed annual total spawner abundance (not density)
vector<lower=0>[n_yrs] dat_esc;
// observed annual spawner age distributions
int<lower=0> dat_age[n_yrs-(age_min+age_skip),AA];
// total catch of adults (no harvest on jacks)
vector[n_yrs+n_fore] dat_harv;                         
}

parameters {
// log intrinsic productivity (units: R/S)
real<lower=-2,upper=10> log_a;
// asymptotic recruitment (units: S)
real<lower=0,upper=1> b;       
// process error SD
real<lower=1e-4,upper=3> sigma_proc;
// process error z-scores
vector[n_yrs-age_min+n_fore] proc_err_z;
// AR(1) coef for proc errors
real<lower=0,upper=1> phi;
// observation error SD
real<lower=1e-4,upper=3> sigma_obs;
// hyper-mean age distribution by cohort
simplex[AA] mu_p;                     
// process error precision of age distribution by cohort
real<lower=0.1,upper=50> prec_p;          
// age distributions by cohort (i.e., brood year)
simplex[AA] pp[n_yrs-age_min+n_fore];                    
// true total spawner abundance in years 1:(age_min+age_skip)
vector<lower=0>[age_min+age_skip] S_tot_init;       
// unknown (missing) spawner abundance obs
vector<lower=0>[max(n_S_NA,1)] dat_esc_NA;    
// true recruits for early missing broods
vector<lower=0>[n_rec_missing] R_early;           
// total catch of adults (no harvest on jacks)
vector<lower=0, upper=1>[n_yrs+n_fore] h_rate;                         
}

transformed parameters {
// carrying capacity
real<lower=0> K;                     
// true total spawner abundance
vector<lower=min(dat_harv)>[n_yrs+n_fore] spawners;           
// true total run size
vector<lower=0>[n_yrs+n_fore] tot_run;           
// age distribution for calendar year returns
simplex[AA] age_props[n_yrs-(age_min+age_skip)];                      
// expected recruits by brood year
vector<lower=0>[n_yrs-age_min+n_fore] R_tot_hat;
// true recruits by brood year
vector<lower=0>[n_yrs-age_min+n_fore] R_tot;           
// process errors
vector[n_yrs-age_min+n_fore] proc_err;
// temp variable: true spawners by age
vector<lower=0>[AA] run;
// recruits by age
vector<lower=0>[AA] Rec[n_yrs-age_min+n_fore];            
// init dummy counter
// vector[2] ret_vec;
// wts for softmax
// simplex[2] wts;            
// total catch of adults (no harvest on jacks)
vector[n_yrs+n_fore] est_harv;                         

if(log_a>0) {
K = log_a / b;
}
else {
K = 0;
}

{
  // init dummy counter
  int cnt;
  cnt = 0;                     
  
  for(t in 1:(n_yrs+n_fore)) {
  if(t <= age_min+age_skip) {
  // 1: early years with no projected recruits
  tot_run[t] = S_tot_init[t];
  }
  else {
  if(t <= age_max) {
  // 2: middle years with incomplete recruits
  // modeled recruits
  for(a in 1:(t-age_min)) {
  run[a] = Rec[t-age_min,a];
  }
  // imputed recruits
  for(a in (t-age_min+1):AA) {
  cnt = cnt + 1;
  run[a] = R_early[cnt];
  }
  }
  else {
  // 3: later years with complete recruits 
  for(a in 1:AA) {
  run[a] = Rec[t-age_min,a];
  }
  }
  // total spawners in years with projected recruits
  tot_run[t] = sum(run);
  // projected calendar-yr age dist
  if(t <= n_yrs) {
  for(a in 1:AA) {
  age_props[t-age_min-age_skip] = run/tot_run[t];
  } 
  }
  }  // end else for middle & late years
  
  // estimated calendar-yr spawners
  // ret_vec[1] = 0;                     
  // ret_vec[2] = tot_run[t] - dat_harv[t];
  // wts = softmax(ret_vec);
  // spawners[t] = wts[1]*1 + wts[2]*ret_vec[2];
  est_harv[t] = h_rate[t] * tot_run[t];
  spawners[t] = tot_run[t] - est_harv[t];
  
  // get projected recruits by age
  if(t <= (n_yrs-age_min+n_fore)) {
  // process errors are AR(1)
  if(t == 1) {
  proc_err[t] = sigma_proc * proc_err_z[t] / sqrt(1 - phi^2);
  }
  else {
  proc_err[t] = phi * proc_err[t-1] + sigma_proc * proc_err_z[t];
  }
  // total recruits
  R_tot_hat[t] = ricker(log_a, b, spawners[t]);
  R_tot[t] = R_tot_hat[t] * exp(proc_err[t]);
  // brood-yr recruits by age
  Rec[t] = R_tot[t] * pp[t];
  }
  
  }  // end loop over t
  
}  // end group for dummy counter
  
}  // end transformed params
  
  model {
  // observed total spawner abundance with missing obs filled in
  vector[n_yrs] dat_esc_aug;     
  
  // Priors
  log_a ~ normal(0,5);
  b ~ normal(0,1);
  phi ~ normal(0,5);
  S_tot_init ~ lognormal(0,10);
  R_early ~ lognormal(0,10);
  if(n_S_NA == 0) {
  dat_esc_NA ~ lognormal(0,10);
  }
  
  // Process models
  // hierarchical model of recruit age distn
  for(t in 1:(n_yrs-age_min+n_fore)) {
  pp[t] ~ dirichlet(mu_p*prec_p);
  }
  // total recruitment by brood year
  // R_tot ~ lognormal(log(R_tot_hat), sigma_proc);
  proc_err_z ~ normal(0,1);
  
  // Observation models
  dat_esc_aug = dat_esc;
  if(n_S_NA > 0) {
  for(t in 1:n_S_NA) {
  dat_esc_aug[which_S_NA[t]] = dat_esc_NA[t];
  }
  }
  // observed total spawners
  dat_esc_aug ~ lognormal(log(spawners), sigma_obs);
  // observed harvest
  dat_harv ~ lognormal(log(est_harv), 0.01);
  // observed age distribution
  for(t in 1:(n_yrs-age_min-age_skip)) {
  dat_age[t] ~ multinomial(age_props[t]);
  }
  }
  
  ", file=file.path(standir, "IPM_RK_AR.stan"))
  ```
  
  ## Ricker with all covars & AR(1)
  
  ```{r stan_RK_cov_AR, eval=FALSE}
  cat("
  
  
  ", file=file.path(standir, "IPM_RK_cov_AR.txt"))
  ```
  
  ## Beverton-Holt with AR(1)
  
  ```{r stan_BH_AR, eval=FALSE}
  cat("
  
  
  ", file=file.path(standir, "IPM_BH_AR.txt"))
  ```
  
  ## Beverton-Holt with all covars & AR(1)
  
  ```{r stan_BH_cov_AR, eval=FALSE}
  cat("
  
  
  ", file=file.path(standir, "IPM_BH_cov_AR.txt"))
  ```
  
  ***
  
  # Fitting the models
  
  Before fitting the model in Stan, we need to specify:
  
  1. the data and indices that go into the model;
  2. the model parameters and states that we want Stan to return;
  3. the MCMC control parameters.
  
  ```{r stan_setup}
  ## 1. Data to pass to Stan
  dat_Stan <- list(
  n_yrs = as.integer(n_yrs),
  AA = as.integer(AA),
  age_min = as.integer(age_min),
  age_max = as.integer(age_max),
  age_skip = as.integer(age_skip),
  n_fore = as.integer(n_fore),
  n_rec_missing = as.integer(((age_max-age_min-age_skip+1)^2 - (age_max-age_min-age_skip+1))/2),
  which_S_NA = which(yr_NA_esc),
  n_S_NA = as.integer(length(which(yr_NA_esc))),
  dat_esc = replace(dat_esc, is.na(dat_esc), 1),
  dat_age = dat_age,
  dat_harv = dat_harv
  )
  if(length(dat_Stan$which_S_NA)==0) { dat_Stan$which_S_NA <- 1 }
  
  ## 2. Model params/states for Stan to return
  ##    These are specific to the process model,
  ##    so we define them in 'par_stan' below.
  
  ## 3. MCMC control params
  mcmc_ctrl <- list(
  chains = 4,
  length = 5e4,
  burn = 2e4,
  thin = 40
  )
  ## total number of MCMC samples after burnin
  mcmc_samp <- mcmc_ctrl$length*mcmc_ctrl$chains/mcmc_ctrl$thin
  ```
  
  ```{r start_timer, include = FALSE}
  ## start timer
  timer_start <- proc.time() 
  ```
  
  ## Models without covariates
  
  Please note that the following code takes ~80 min to run on a quad-core machine with 3.5 GHz Intel processors.
  
  ```{r stan_inits}
  ## empty list for fits
  n_mods <- 4
  mod_fits <- vector("list", n_mods)
  
  ## function to create Stan inits
  init_Stan <- function() {
  for(i in names(dat_Stan)) assign(i, dat_Stan[[i]])
  if(n_S_NA > 0) {
  S_tot_obs_noNA <- dat_esc
  dat_esc[which_S_NA] <- NA
  }
  mu_p <- colSums(dat_age)/sum(dat_age)
  R_tot <- rpois(n_yrs-age_min+n_fore,median(dat_esc[-which_S_NA]))
  inits <- list(
  log_a = runif(1,2,3),
  b = 1/rpois(1,median(dat_esc[-which_S_NA])),
  sigma_proc = runif(1,0.1,0.5),
  proc_err_z = rnorm(n_yrs-age_min+n_fore),
  phi = runif(1,0,0.9),
  sigma_obs = runif(1,0.1,0.5), 
  mu_p = mu_p,
  prec_p = runif(1,1,2),
  pp = matrix(mu_p,n_yrs-age_min+n_fore,AA,byrow=TRUE),
  S_tot_init = rep(max(S_tot_obs_noNA), age_min+age_skip),
  dat_esc_NA = array(rep(mean(dat_esc[-which_S_NA]), max(n_S_NA, 1)), dim=max(n_S_NA, 1)),
  R_tot = R_tot,
  R_early = rpois(((age_max-age_min-age_skip+1)^2-(age_max-age_min-age_skip+1))/2,1000)
  )
  return(inits)
  }
  ```
  
  ### Ricker
  
  ```{r fit_Ricker_base, message = FALSE, warning = TRUE, cache = FALSE}
  ## params/states for Stan to return
  par_Stan <- c("log_a", "b", "K",
  "pp", "mu_p", "prec_p",
  "sigma_proc", "sigma_obs",
  "R_tot", "Rec", "spawners")
  
  ## fit the model in Stan & store results
  mod_fits[[1]] <- stan(file = file.path(standir, "IPM_RK_AR.stan"),
  data = dat_Stan, 
  init = init_Stan, 
  pars = par_Stan,
  control=list(adapt_delta=0.99, max_treedepth=13),
  chains = 2, iter = 6000, warmup = 1000, thin = 20)
  ```
  
  ```{r save_model, eval=TRUE}
  best_fit <- mod_fits[[1]]
  save(best_fit, file = file.path(analdir, "best_fit.rda"))
  ```
  
  
  #### Convergence checks
  
  ```{r diag_Ricker, eval=FALSE}
  par_conv <- c("alpha","beta","sigma_r","sigma_s","pi_tau",paste0("pi_eta[",seq(AA-1),"]"))
  
  ## Gelman-Rubin
  gelman.diag(mod_fits[[1]][,par_conv])
  
  ## autocorrelation
  t(round(autocorr.diag(mod_fits[[1]][,par_conv],
  lags = seq(mcmc_ctrl$thin, 4*mcmc_ctrl$thin, mcmc_ctrl$thin),
  relative=FALSE), 2))
  ```
  
  ```{r stan_diag}
  #------------------
  # Stan diagnostics
  #------------------
  
  # summary of Stan fit
  show(mod_fits[[1]])
  
  # plot summary of Stan fit
  #plot(mod_fits[[1]])
  
  # HMC sampler stats: all chains combined
  #summary(do.call(rbind, args = get_sampler_params(mod_fits[[1]], inc_warmup = FALSE)), digits = 2)
  
  # HMC sampler stats: each chain separately
  #lapply(get_sampler_params(mod_fits[[1]], inc_warmup = FALSE), summary, digits = 2)
  
  # number of post-sampler divergent transitions
  #count_divergences(mod_fits[[1]])
  
  #hist_treedepth(mod_fits[[1]])
  
  
  #mod.list <- As.mcmc.list(mod_fits[[1]])
  ```
  
  ```{r, eval=FALSE}
  Rec <- extract(mod_fits[[1]], pars = "Rec", permuted = TRUE, inc_warmup = TRUE)$Rec
  R_tot <- unlist2(extract(mod_fits[[1]], pars = "R_tot", permuted = TRUE, inc_warmup = TRUE))
  pp <- extract(mod_fits[[1]], pars = "pp", permuted = TRUE, inc_warmup = TRUE)$Rec
  
  m1 <- as.array(mod_fits[[1]])
  #dev.new()
  par(mfrow=c(4,1), mai=c(0.6,0.6,0.1,0.1))
  apply(m1[,,"log_a"],2,acf)
  
  m2 <- as.matrix(mod_fits[[1]])
  #dev.new()
  pairs(mod_fits[[1]], pars=c("b","log_a","K"))
  
  pairs(mod_fits[[1]], pars=c("mu_p","prec_p"))
  
  pairs(mod_fits[[1]], pars=c("sigma_proc","sigma_obs"))
  
  traceplot(mod_fits[[1]], pars=c("sigma_proc","sigma_obs"))
  
  #traceplot(mod_fits[[1]], pars=colnames(m2)[grep("Rec",colnames(m2))])
  
  get_posterior_mean(mod_fits[[1]], par=c("K","log_a","sigma_proc","sigma_obs"))
  
  ```
  
  