---
title: Appendix S1. Detailed instructions for replicating analyses and reproducing
  results.
output:
  html_document:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

***

This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

***

## Model diagnostics

Here is a histogram of the Gelman & Rubin statistics $(R_{hat})$ for the estimated parameters. Recall that we set an upper threshold of `r Rhat_thresh`, so values larger than that deserve some additional inspection.

```{r model_diagnostics}
## Rhat values for all parameters
rh <- mod_fit$BUGSoutput$summary[,"Rhat"]
## histogram of Rhat values for all parameters
par(mai=c(0.9,0.9,0.3,0.1))
hist(rh, breaks=seq(1,ceiling(max(rh)/0.01)*0.01,by=0.01),main="",
     col=rgb(0,0,255,alpha=50,maxColorValue=255),border="blue3",xlab=expression(italic(R[hat])))
## Rhat values > threshold
bad_Rhat <- rh[rh>Rhat_thresh]
## prop of params with Rhat > threshold
round(length(bad_Rhat)/length(rh),3)
## param names
par_names <- sub("\\[.*","",names(bad_Rhat))
## number of Rhat > threshold by param name
table(par_names)
## index values for offenders
idx <- as.integer(sub("(^.*\\[)([0-9]{1,3})(.*)","\\2",names(bad_Rhat)))
## data frame of offenders
(df <- data.frame(par=par_names, index=idx))
```

The convergence statistics indicate that some of the elements in $p$ for the estimated proportions of the youngest and oldest age classes (i.e., 3 and 8, respectively) did not converge to our desired threshold. However, there is very little data to inform those parameters, so we should not be too concerned.

## Main results

Here is a table of summary statistics for some of the model parameters.

```{r tbl_summary_stats}
tbl_smry <- mod_fit$BUGSoutput$summary[c("E_Rkr_a","beta",
                                         paste0("gamma[",seq(n_cov),"]"),
                                         "sigma_r","sigma_s"),
                                       c("mean","sd","2.5%","50%","97.5%")]
rownames(tbl_smry)[seq(n_cov)+2] <- colnames(dat_cvrs)                                 
print(tbl_smry,digits=3,quote=FALSE,justify="right")
```

### Spawner-recruit relationship

Here is the relationship between spawner and subsequent recruits (a), assuming mean values for all covariates. Gray lines show the median relationship for each of the `r n_yrs` years based on $a_t$. Note that for plotting purposes only in (b) and (c), the density in the largest bin for each parameter contains counts for all values greater or equal to that. Vertical arrows under the x-axes in (b) and (c) indicate the 2.5^th^, 50^th^, and 97.5^th^ percentiles.

```{r plot_S_R, fig.width=8, fig.height=5, fig.pos="placeHere"}
layout(matrix(c(1,1,2,3),2,2),c(3,2),c(1,1))
CI_vec <- c(0.025,0.5,0.975)
offSet <- 0.06

## posterior of spawners
sDat <- apply(mod_fit$BUGSoutput$sims.list$Sp,2,quantile,CI_vec)
sDat <- sDat[,1:(n_yrs-age_min+n_fore)]
## posterior of recruits
rDat <- exp(apply(mod_fit$BUGSoutput$sims.list$tot_ln_Rec,2,quantile,CI_vec))
## median values for a & b
aa <- apply(mod_fit$BUGSoutput$sims.list$ln_Rkr_a,2,median)
bb <- apply(mod_fit$BUGSoutput$sims.list$beta,2,median)

## empty plot space for spawner-recruit relationships
dd <- 3000
yM <- Re2prec(max(rDat),"ceiling",dd)
yM <- 30000
xM <- Re2prec(max(sDat),"ceiling",dd)
par(mai=c(0.8,0.8,0.1,0.1), omi=c(0,0,0,0))
plot(sDat[2,],rDat[2,], xlim=c(0,xM), ylim=c(0,yM), pch=16, col="blue3", type="n",
	 xaxs="i", yaxs="i", ylab="Recruits (1000s)", xlab="Spawners (1000s)", cex.lab=1.2,
	 xaxt="n", yaxt="n")
axis(1, at=seq(0,xM,dd*2), labels=seq(0,xM,dd*2)/1000)
axis(2, at=seq(0,yM,dd*2), labels=seq(0,yM,dd*2)/1000)
for(i in 1:length(aa)) { lines(seq(xM)*exp(aa[i]-bb*seq(xM)), col="darkgray") }
## add S-R estimates and medians
abline(a=0,b=1,lty="dashed")
nCB <- n_yrs-age_max
points(sDat[2,1:nCB],rDat[2,1:nCB], xlim=c(0,xM), ylim=c(0,yM), pch=16, col="blue3")
segments(sDat[2,1:nCB],rDat[1,1:nCB],sDat[2,1:nCB],rDat[3,1:nCB], col="blue3")
segments(sDat[1,1:nCB],rDat[2,1:nCB],sDat[3,1:nCB],rDat[2,1:nCB], col="blue3")
nTB <- dim(sDat)[2]
clr <- rgb(100, 0, 200, alpha=seq(200,100,length.out=age_max-age_min+n_fore), maxColorValue=255)
segments(sDat[2,(nCB+1):nTB],rDat[1,(nCB+1):nTB],sDat[2,(nCB+1):nTB],rDat[3,(nCB+1):nTB], col=clr)
segments(sDat[1,(nCB+1):nTB],rDat[2,(nCB+1):nTB],sDat[3,(nCB+1):nTB],rDat[2,(nCB+1):nTB], col=clr)
points(sDat[2,(nCB+1):nTB],rDat[2,(nCB+1):nTB],
       xlim=c(0,xM), ylim=c(0,yM), pch=16, col=clr)
text(x=par()$usr[1]+par()$pin[2]/par()$pin[1]*offSet*diff(par()$usr[1:2]),
	 y=par()$usr[4]-offSet*diff(par()$usr[3:4]),"(a)")

## posterior for alpha
clr <- rgb(0, 0, 255, alpha = 50, maxColorValue = 255)
par(mai=c(0.8,0.4,0.3,0.1))
## Ricker a
R_alpha_est <- mod_fit$BUGSoutput$sims.list$alpha
alphaCI <- quantile(R_alpha_est,c(0.025,0.5,0.975))
R_alpha_est[R_alpha_est>9] <- 9
hist(R_alpha_est,freq=FALSE,xlab="",main="",breaks=seq(0,9,0.2),
     col=clr, border="blue3", ylab="", cex.lab=1.2, yaxt="n")
aHt <- (par()$usr[4]-par()$usr[3])/12
arrows(alphaCI,par()$usr[3],alphaCI,par()$usr[3]-aHt,
       code=1,length=0.05,xpd=NA,col="blue3",lwd=1.5)
mtext(expression(Instrinsic~productivity~(alpha)), 1, line=3, cex=1)
text(x=par()$usr[1]+par()$pin[2]/par()$pin[1]*offSet*diff(par()$usr[1:2]),
	 y=par()$usr[4]-offSet*diff(par()$usr[3:4]),"(b)")

## posterior for K
par(mai=c(0.8,0.4,0.3,0.1))
aa <- matrix(mod_fit$BUGSoutput$sims.array[,,"E_Rkr_a"],ncol=1)
bb <- matrix(mod_fit$BUGSoutput$sims.array[,,"beta"],ncol=1)
R_b_est <- aa/bb
R_b_est <- R_b_est[R_b_est > 0]
R_b_CI <- quantile(R_b_est,c(0.025,0.5,0.975))
R_b_est[R_b_est>2e4] <- 2e4
brks <- seq(Re2prec(min(R_b_est),"floor",2000),2e4,length.out=length(seq(0,9,0.2)))
hist(R_b_est, freq=FALSE, breaks=brks, col=clr, border="blue3",
	 xlab="", xaxt="n", yaxt="n",
	 main="", ylab="", cex.lab=1.2)
axis(1, at=seq(Re2prec(min(R_b_est),"floor",2000),1.8e4,4000))
aHt <- (par()$usr[4]-par()$usr[3])/12
arrows(R_b_CI,par()$usr[3],R_b_CI,par()$usr[3]-aHt,
       code=1,length=0.05,xpd=NA,col="blue3",lwd=1.5)
mtext(expression(Carrying~capacity~(italic(K))), 1, line=3, cex=1)
text(x=par()$usr[1]+par()$pin[2]/par()$pin[1]*offSet*diff(par()$usr[1:2]),
	 y=par()$usr[4]-offSet*diff(par()$usr[3:4]),"(c)")
```

### Covariate effects

Here are time series plots of the covariates (a-e) and histograms of their effects on productivity (f-j).

```{r plot_cov_effects, fig.width=8, fig.height=10, fig.pos="placeHere", warnings=FALSE, messages=FALSE, eval=FALSE}
clr <- rgb(0, 0, 255, alpha = 50, maxColorValue = 255)
offSet <- 0.07
c_est <- mod_fit$BUGSoutput$sims.list$gamma
par(mfrow=c(n_cov,2), mai=c(0.4,0.2,0.1,0.1), omi=c(0.2,0.5,0,0))
ylN <- floor(min(c_est)*10)/10
ylM <- ceiling(max(c_est)*10)/10
brks <- seq(ylN,ylM,length.out=diff(c(ylN,ylM))*40+1)
cov_names <- c(expression(paste("Max flow (",m^3," ",s^{-1},")")),
               expression(paste("Min flow (",m^3," ",s^{-1},")")),
               "NPGO",
               "STI (day of year)",
               expression(paste("H releases (",10^3,")")))
t_idx <- seq(yr_frst,length.out=n_yrs-age_min+n_fore)
for(i in 1:n_cov) {
  if(i==5) {dat_cvrs[,i] <- dat_cvrs[,i]/1000}
	## plot covar ts
	plot(t_idx, dat_cvrs[seq(length(t_idx)),i], xlab="", ylab="",
		 main="", cex.axis=1.2, pch=16, col="blue3", type="o", bty="L")
	text(x=par()$usr[1]+par()$pin[2]/par()$pin[1]*offSet*diff(par()$usr[1:2]),
	     y=par()$usr[4]-offSet*diff(par()$usr[3:4]),paste0("(",letters[i],")"), cex=1.2)
	mtext(side=2, cov_names[i], line=3, cex=1.2)
	if(i==n_cov) { mtext(side=1,"Brood year", line=3) }
	## plot covar effect
	hist(c_est[,i],
	     freq=FALSE,breaks=brks,col=clr,border="blue3",
	     xlab="", yaxt="n", main="", ylab="", cex.axis=1.2)
	c_CI <- quantile(c_est[,i],c(0.025,0.5,0.975))
	aHt <- (par()$usr[4]-par()$usr[3])/20
	arrows(c_CI,par()$usr[3]-0.005,c_CI,par()$usr[3]-aHt,
	       code=1,length=0.05,xpd=NA,col="blue3",lwd=1.5)
	abline(v=0, lty="dashed")
	text(x=par()$usr[1]+par()$pin[2]/par()$pin[1]*offSet*diff(par()$usr[1:2]), cex=1.2,
	     y=par()$usr[4]-offSet*diff(par()$usr[3:4]),paste0("(",letters[i+n_cov],")"))
	if(i==n_cov) { mtext(side=1,"Effect size", line=3) }
}
```

### Management reference points

Here are a number of management reference points.

```{r ref_pts}
# abbreviations for ref points
ref_names <- c("MSY","Smsy","Umsy","Umax")
# proportions of MSY to consider
yld_prop <- c(0.75,0.85,0.95)
aa <- matrix(mod_fit$BUGSoutput$sims.array[,,"E_Rkr_a"],ncol=1)
alpha <- matrix(mod_fit$BUGSoutput$sims.array[,,"alpha"],ncol=1)
mcmc <- length(aa)
# empty matrix for ref pts
ref_pts <- matrix(NA,mcmc,length(ref_names))
colnames(ref_pts) <- ref_names
# spawner series for optimal yield profile
SS <- seq(100,1e4,100)
# empty matrix for optimal yield profiles
OYP <- matrix(0,length(SS),length(yld_prop))
for(i in 1:mcmc) {
	# spawners at MSY
	ref_pts[i,"Smsy"] <- (1 - lambert_W0(exp(1-aa[i]))) / bb[i]
	# MSY
	ref_pts[i,"MSY"] <- ref_pts[i,"Smsy"]*((exp(aa[i]-bb[i]*ref_pts[i,"Smsy"])) - 1)
	# harvest rate at MSY
	ref_pts[i,"Umsy"] <- (1 - lambert_W0(exp(1-aa[i])))
	# max harvest rate
	ref_pts[i,"Umax"] <- 1 - 1/alpha[i]
	# yield over varying S
	yield <- SS*(exp(aa[i]-bb[i]*SS) - 1)
	for(j in 1:length(yld_prop)) {
		OYP[,j] <- OYP[,j] + 1*(yield > yld_prop[j]*ref_pts[i,"MSY"])
	}
}
OYP <- OYP/mcmc

## Prob of overfishing
hh <- seq(100)
Pr_over <- cbind(hh,hh,hh)
colnames(Pr_over) <- c("Umsy75","Umsy","Umax")
for(i in hh) {
  Pr_over[i,"Umsy75"] <- sum(ref_pts[,"Umsy"]*0.75 < i/100)/mcmc_samp
  Pr_over[i,"Umsy"] <- sum(ref_pts[,"Umsy"] < i/100)/mcmc_samp
  Pr_over[i,"Umax"] <- sum(ref_pts[,"Umax"] < i/100)/mcmc_samp
}

## posterior exploitation rate & spawner abundance
aer <- Sp_ts <- mod_fit$BUGSoutput$sims.list$Sp[,1:n_yrs]
for(i in 1:n_yrs) {
	aer[,i] <- dat_harv[i] / (dat_harv[i] + Sp_ts[,i]) 
}
```

These are plots of (a) the probability that a given number of spawners produce average yields exceeding X% of MSY (i.e, optimal yield profiles); and (b) the cumulative probabilty of overfishing the population, based on harvest rates equal to those at 75% of MSY $(U_{\text{M75}})$, MSY $(U_{\text{MSY}})$, and the maximum $(U_{\text{Max}})$. The probability of exceeding $U_{\text{Max}}$ indicates the risk that offspring will not replace their parents, which, if sustained, will lead to eventual extinction. The histograms above (a) and (b) are distributions of the posterior estimates for the number of spawners and harvest rates, respectively.

```{r plot_ref_pts, fig.width=8, fig.height=4, warning=FALSE}
layout(matrix(c(2,1,4,3),2,2),heights=c(1,5))

## OYP
par(mai=c(0.9,0.9,0,0), omi=c(0,0,0.1,0.1))
x_lp <- yld_prop
for(i in 1:length(x_lp)) {
	x_lp[i] <- SS[max(which(OYP[,i] == max(OYP[,i]) | abs(OYP[,i] - (yld_prop[i]-0.3)) <= 0.05))]
}
matplot(SS, OYP, type="l", lty="solid", las=1, col=c("slateblue","blue","darkblue"), lwd=2,
		xlab="Spawners", ylab="Probability of X% of MSY", cex.lab=1.2,
		main="", ylim=c(0,1))
points(x=x_lp, y=yld_prop-0.3, pch=21, cex=3.5, col="white", bg="white")
text(x=x_lp, y=yld_prop-0.3, paste0(yld_prop*100,"%"),
	 col=c("slateblue","blue","darkblue"), cex=0.7)
text(x=par()$usr[1]+par()$pin[2]/par()$pin[1]*offSet*diff(par()$usr[1:2]),
     y=par()$usr[4]-offSet*diff(par()$usr[3:4]),"(a)")
## posterior spawner abundance over all years
par(mai=c(0,0.9,0.05,0))
hist(Sp_ts[Sp_ts<1e4], col=clr, border="blue3", breaks=40,
	  main="", yaxs="i", xaxt="n",yaxt="n",ylab="")

## prob of overfishing
par(mai=c(0.9,0.9,0,0))
matplot(Pr_over, type="l", las=1, lwd=2, lty="solid", col=c("slateblue","blue","darkblue"),
        ylab="Probability of overfishing", cex.lab=1.2,
        xlab="Harvest rate", xaxt="n")
axis(1,seq(0,100,20),seq(0,100,20)/100)
x_lp <- c(0,0,0)
for(i in 1:length(x_lp)) {
	x_lp[i] <- max(which(abs(Pr_over[,i] - 0.5) <= 0.05))
}
points(x=x_lp, y=rep(0.5,3), pch=21, cex=4, col="white", bg="white")
text(x=x_lp, y=0.5, expression(U[M75], U[MSY], U[Max]),
	 col=c("slateblue","blue","darkblue"), cex=0.8)
text(x=par()$usr[1]+par()$pin[2]/par()$pin[1]*offSet*diff(par()$usr[1:2]),
     y=par()$usr[4]-offSet*diff(par()$usr[3:4]),"(b)")
## posterior harvest rates over all years
par(mai=c(0,0.9,0.05,0))
hist(aer, col=clr, border="blue3", breaks=seq(0,40)/40,
	  main="", yaxs="i", xaxt="n",yaxt="n",ylab="")
```

### Total population size

Here is our estimate of the total run size (i.e., catch + escapement) over time, which includes a forecast for 2016. The black points are the data, the blue line is the median posterior estimate, and the shaded region is the 95% credible interval. Note that the y-axis is on a log scale.

```{r plot_run_size, fig.width=6, fig.height=4.5, fig.pos="placeHere"}
pDat <- apply(mod_fit$BUGSoutput$sims.list$Sp,2,quantile,probs=CI_vec)
pDat <- pDat + matrix(dat_harv,length(CI_vec),n_yrs+n_fore,byrow=TRUE)
t_idx_f <- seq(yr_frst,length.out=n_yrs+n_fore)
ypMin <- min(pDat)
ypMax <- max(pDat)
par(mai=c(0.8,0.8,0.1,0.1), omi=c(0.5,0.2,0.1,0.2))
plot(t_idx_f,pDat[3,], ylim=c(ypMin,ypMax), type="n", log="y", xaxt="n", yaxt="n",
	 xlab="Year", ylab="Catch + escapement", main="", cex.lab=1.2)
polygon(c(t_idx_f,rev(t_idx_f)),c(pDat[3,],rev(pDat[1,])), col=clr, border=NA)
lines(t_idx_f, pDat[2,], col="blue3", lwd=2)
points(t_idx_f, exp(ln_dat_esc)+dat_harv, pch=16, cex=1)
axis(1,at=seq(1980,2015,5))
axis(2,at=c(4000,8000,16000))
```

Here are several percentiles for the 2016 forecast for the total run size (i.e., catch + escapement).

```{r forecast_2016}
data.frame(forecast=round(quantile(mod_fit$BUGSoutput$sims.list$Sp[,n_yrs+n_fore],
                                   probs=c(0.025,0.25,0.5,0.75,0.975))))
```

### Correlation of innovations & SAR

Here is our estimate of the correlation between the model innovations (i.e, process residuals) and the smolt-to-adult return rate (i.e., logit transformed proportion of smolts that survive to adulthood) of hatchery steelhead from the Skagit R (Neala Kendall, WDFW, Olympia, WA, unpublished data). This gives an indication as to whether the unexplained variance in the productivity of wild fish might be related to the marine environment.

```{r SAR_correl, eval=FALSE}
## get SAR data
dat_SAR <- read_csv(paste0(ex_url,"skagit_sthd_hsar.csv"))
## match brood yr to smolt outmigration yr
dat_SAR$br_yr <- dat_SAR$out_yr - 2
## get innov
innov <- t(mod_fit$BUGSoutput$sims.list$res_ln_Rec)
## trim to same brood yrs as SAR data
innov <- innov[t_idx %in% dat_SAR$br_yr,]
## compute correlation over all mcmc samples
cor_vec <- apply(innov,2,function(x) { cor(qlogis(dat_SAR$SAR),x) })
## print median -/+ 95% credible interval
print(quantile(cor_vec,CI_vec), digits=2)
```

The distribution of correlation coefficients has a median of 0.29, with a 95% credible interval of (0.032,0.52), which would suggest that the productivity of wild steelhead from the Skagit is also affected by the marine environment.

***

## Additional results

The following results are not reported in the main manuscript, but are presented here as additional examples of other model estimates.

### Spawners over time

Here is the estimate of the number of spawners over time. The black points are the data, the blue line is the median posterior estimate, and the shaded region is the 95% credible interval. Note that there are no estimates of spawners in 1996 & 1997.

```{r plot_spawners, fig.width=6, fig.height=4, fig.pos="placeHere"}
pDat <- apply(mod_fit$BUGSoutput$sims.list$Sp,2,quantile,CI_vec)
ypMin <- min(pDat[,1:n_yrs])
ypMax <- max(pDat[,1:n_yrs])
t_idx_T <- seq(yr_frst,length.out=n_yrs)
par(mai=c(0.8,0.8,0.1,0.1), omi=c(0,0.2,0.1,0.2))
plot(t_idx_T,pDat[3,1:n_yrs], ylim=c(ypMin,ypMax), type="n", log="y", xaxt="n", yaxt="n",
	 xlab="Year", ylab="Spawners", main="", cex.lab=1.2)
polygon(c(t_idx_T,rev(t_idx_T)),c(pDat[3,1:n_yrs],rev(pDat[1,1:n_yrs])), col=clr, border=NA)
lines(t_idx_T, pDat[2,1:n_yrs], col="blue3", lwd=2)
points(seq(yr_frst,length.out=n_yrs+n_fore), exp(ln_dat_esc), pch=16, cex=1)
axis(1,at=seq(1980,2015,5))
axis(2,at=c(3000,6000,12000))
```

### Recruits over time

Here are the estimated total number of recruits by brood year (note that the y-axis is on a log scale). Again the uncertainty increases in recent years because fewer complete age classes have been observed.

```{r plot_total_recruits, fig.width=6, fig.height=4, fig.pos="placeHere"}
CI_vec <- c(0.025,0.5,0.975)
pDat <- apply(mod_fit$BUGSoutput$sims.list$Rec,c(1,2),sum)
pDat <- apply(apply(pDat,2,sort),2,function(x) { x[mcmc_samp*CI_vec] })
ypMin <- min(pDat)
ypMax <- max(pDat)
t_idx_a <- seq(yr_frst,length.out=n_yrs-age_min+n_fore)
par(mai=c(0.8,0.8,0.1,0.1), omi=c(0,0.2,0.1,0.2))
plot(t_idx_a,pDat[3,], ylim=c(ypMin,ypMax), type="n", log="y", yaxt="n",
	 xlab="Brood year", ylab="Recruits", main="", cex.lab=1.2)
axis(2,at=c(3000,9000,27000))
polygon(c(t_idx_a,rev(t_idx_a)),c(pDat[3,],rev(pDat[1,])), col=clr, border=NA)
lines(t_idx_a, pDat[2,], col="blue3", lwd=2)
```

### Recruits per spawner

Here is the time series of estimated recruits-per-spawner. Values above (below) the dashed line at zero indicate positive (negative) population growth.

```{r plot_R_per_S, fig.width=6, fig.height=4, fig.pos="placeHere"}
pDat <- apply(mod_fit$BUGSoutput$sims.list$ln_RS,2,sort)
pDat <- apply(pDat,2,function(x) { x[mcmc_samp*CI_vec] })
pDat[2,] <- apply(mod_fit$BUGSoutput$sims.list$ln_RS,2,median)
ypMin <- min(pDat)
ypMax <- max(pDat)
par(mai=c(0.8,0.8,0.1,0.1), omi=c(0,0.2,0.1,0.2))
plot(t_idx_a,pDat[3,], ylim=c(ypMin,ypMax), type="n", #log="y",
	 xlab="Brood year", ylab="ln(R/S)", main="", cex.lab=1.2)
abline(h=0, lty="dashed")
polygon(c(t_idx_a,rev(t_idx_a)),c(pDat[3,],rev(pDat[1,])), col=clr, border=NA)
lines(t_idx_a, pDat[2,], col="blue3", lwd=2)
```

### Age composition

Here are time series of the estimated proportions of each age class by brood year (cohort).

```{r plot_age_comp, fig.width=6, fig.height=4, fig.pos="placeHere"}
par(mai=c(0.8,0.8,0.1,0.1), omi=c(0,0.1,0.2,0.2))
clr <- rgb(0, 0, 255, alpha = 40, maxColorValue = 255)
age_est <- t(apply(apply(mod_fit$BUGSoutput$sims.list$pi_vec,c(3,2),mean),2,cumsum))
nRec <- n_yrs-age_min
plot(t_idx_a, rep(1,nRec+n_fore), ylab="Proportion", xlab="Brood year", ylim=c(0,1), las=1,
     xaxs="i", yaxs="i", type="n", lty="solid", col="blue3", cex.lab=1.2)
for(i in c(1,2,3,4,6)) {
	polygon(c(t_idx_a,rev(t_idx_a)),c(age_est[,i],rep(0,nRec+n_fore)), col=clr, border=NA)
	}
lbl <- apply(cbind(c(0,age_est[nRec+n_fore,-A]),age_est[nRec+n_fore,]),1,mean)
text(par()$usr[2],par()$usr[4]*1.05,"Age", xpd=NA, pos=4, offset=0.05, col="black", cex=0.8)
text(par()$usr[2],lbl[1:4],seq(3,6), xpd=NA, pos=4, col="black", cex=0.7)
text(par()$usr[2],lbl[5],"7&8", xpd=NA, pos=4, offset=0.15, col="black", cex=0.7)
```

### Recruits by age class

Here are the estimated number of recruits by brood year and age. Note that the uncertainty increases in recent years as fewer complete age classes have been observed.

```{r plot_recruits_by_age, fig.width=6, fig.height=6, fig.pos="placeHere"}
CI_vec <- c(0.05,0.5,0.95)
par(mfrow=c(A,1), mai=c(0.1,0.1,0.05,0.1), omi=c(0.5,0.5,0.1,0))
t_idx_R <- seq(yr_frst,length.out=nRec+n_fore)
pltTT <- seq(min(round(t_idx_R/5,0)*5),max(round(t_idx_R/5,0)*5),5)
for(i in rev(1:A)) {
	pDat <- apply(mod_fit$BUGSoutput$sims.list$Rec[,,i],2,sort)
	pDat <- apply(pDat,2,function(x) { x[mcmc_samp*CI_vec] })/100
	dd <- ifelse(max(pDat)<20,1,10)
	ypMax <- Re2prec(max(pDat),prec=dd)
	while(ypMax %% 3 != 0) { ypMax <- ypMax + dd }
	plot(t_idx_R,pDat[3,], xlim=c(yr_frst+1,yr_last-n_fore-2), ylim=c(0.001,ypMax),
	     type="n", xaxt="n", yaxt="n", xlab="", ylab="", main="", las=1)
	polygon(c(t_idx_R,rev(t_idx_R)),c(pDat[3,],rev(pDat[1,])), col=clr, border=NA)
	lines(t_idx_R, pDat[2,], col="blue3", lwd=2)
	aHt <- (par()$usr[4]-par()$usr[3])/7
	ttl <- paste("Age-",i+age_min-1,sep="")
	text(t_idx_R[1]-0, par()$usr[4]-aHt, ttl, pos=4, cex=0.9)
	axis(2,seq(0,ypMax,length.out=4),las=1,cex=0.9)
	if(i!=1) {axis(1,at=pltTT,labels=FALSE)} else {axis(1,at=pltTT)}
}
mtext("Recruits (100s)", 2, line=2, outer=TRUE, cex=1.2)
mtext("Year", 1, line=2.5, outer=TRUE, cex=1.2)
```

### Time-varying productivity

Here is the time series of the time-varying productivity ($a_t$), which includes the cumulative effects of the `r n_cov` covariates.

```{r plot_prod_over_time, fig.width=6, fig.height=4, fig.pos="placeHere"}
pDat <- apply(mod_fit$BUGSoutput$sims.list$ln_Rkr_a,2,quantile,CI_vec)
ypMin <- min(pDat)
ypMax <- max(pDat)
par(mai=c(0.8,0.8,0.1,0.1), omi=c(0,0.2,0.1,0.2))
plot(t_idx_a,pDat[3,], ylim=c(ypMin,ypMax), type="n", #log="y",
	 xlab="Brood year", ylab="Productivity", main="", cex.lab=1.2)
polygon(c(t_idx_a,rev(t_idx_a)),c(pDat[3,],rev(pDat[1,])), col=clr, border=NA)
lines(t_idx_a, pDat[2,], col="blue3", lwd=2)
```

### Innovations

Here is the time series of the so-called "innovations", which are the residuals from the process model. They give some indication of population productivity after accounting for the effects of density dependence.

```{r plot_innovations, fig.width=6, fig.height=4, fig.pos="placeHere"}
pDat <- apply(mod_fit$BUGSoutput$sims.list$res_ln_Rec,2,quantile,CI_vec)
ypMin <- min(pDat)
ypMax <- max(pDat)
par(mai=c(0.8,0.8,0.1,0.1), omi=c(0,0.2,0.1,0.2))
plot(t_idx_a,pDat[3,], ylim=c(ypMin,ypMax), type="n", #log="y",
	 xlab="Brood year", ylab="Innovations", main="", cex.lab=1.2)
abline(h=0, lty="dashed")
polygon(c(t_idx_a,rev(t_idx_a)),c(pDat[3,],rev(pDat[1,])), col=clr, border=NA)
lines(t_idx_a, pDat[2,], col="blue3", lwd=2)
```
